\thetitle

### 2020-07-01 writing/challenges.tex
organize the challenge section

Realizing these ideas into a robust experiment requires surpassing three challenges.  First,
gradual typing is a heterogeneous landscape. Herein, we focus on three
points of the landscape: the natural approach of Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010},
the transient one of Reticulated
Python~\cite{vsc-dls-2019,vss-popl-2017,vksb-dls-2014} and  the optional
gradual typing of the likes of Typescript. Each of these points comes with
a distinct blaming strategy and together they represent all proposed
strategies so far. In addition, instead of experimenting with the three
approaches in isolation, we have implemented transient and optional
gradual typing as variants of  Typed Racket to seize the opportunity to
compare them.  However, implementations on top of the same platform are
not enough for an apples-to-apples comparison. The three approaches differ
not only on their blaming strategies but also on the type checks they
perform. To untangle the two, we supplement our experiment with two extra
modes. The \emph{vanilla} mode  ignores blame and instead uses information
from dynamic type error messages to select how to proceed with migration. Hence it
separates the effect of blame on the experiment from that of the type
checks. The \emph{control} mode randomly selects which part of the program to migrate
and serves as the null hypothesis control for the overall methodology.

The second challenge stems from the lack of a sufficiently large available
corpus of type errors for a meaningful experiment. Similar to
~\citet{lksfd-popl-2020}, we cover this gap by injecting bugs in
typed programs  with mutation analysis~\cite{lipton1971fault,
demillo1978hints, jia2011analysis}. This results into a large 
number of buggy programs whose mix-typed versions are suitable subjects
for our experiment. However, mutation analysis
traditionally  aims to inject bugs that are sophisticated enough to
challenge extensive test suites. As a consequence, the ill-typed mutants we care about here are
usually deemed \emph{incompetent} and mutation analysis frameworks are
fine-tuned to avoid them. To overcome this problem, we design a
set of mutators specific for type errors inspired by common errors when
programming with in Typed Racket.  These type errors cover
a wide range of Typed Racket's type system and moreover cause reliably runtime
type errors  in mix-typed versions of mutants.

The final challenge deals with the fact that bugs in mix-typed programs
occur not only in the code but also in type
annotations~\cite{sta-nt-base-types, incorrect-ts, wmwz-ecoop-2017}. Thus
our experiment needs to also explore the relation between blame and the
location of bugs in case of wrong type annotations. Similarly to the previous challenge, we
deal with this one by designing a set of mutators that ``break'' type
annotations but leave programs otherwise intact.

To sum up, the maim {\bf contribution} of this paper is an empirical  methodology for
evaluating blame for gradual type systems. The methodology applies across the
landscape of gradual typing and allows to compare its different points. 
  As a result, this paper also contributes the first comparative
empirical study of natural, transient and optional gradual typing
with respect to blame. [A couple of sentences that summarize the results
her.]

The remainder of the paper is organized as follows.
Section~\ref{sec:landscape} revisits the landscape of gradual typing from
the perspective of the different checking and blame strategies.
Section~\ref{sec:ideas} presents in detail the main ideas behind our
methodology and section~\ref{sec:experiment} explains the specifics of our
experiment. Section~\ref{sec:results} gathers our results and
section~\ref{sec:discussion} discusses their implications. The paper
concludes with a survey of related work in section~\ref{sec:related} and 
a few final remarks in section~\ref{sec:conclusion}.

>>>>

The most serious challenge concerns the nature of this research. Unlike ordinary
work on programming languages, the question seems to call for research similar
to the one in the human-computer interaction area. At the same time, a
significant evaluation result demands a good number of cases yielding a serious
amount of data. But, as~\citet{lksfd-popl-2020} already demonstrated, the way
around this dilemma is to postulate the equivalent of a {\it home economicus\/}
for programming languages---the {\em rational programmer\/}---and to implement
this rational programmer as an algorithm. As indicated in the introduction, in
the context of gradual typing a rationally acting programmer translates the
Wadler-Findler slogan into a debugging method, searching for the source of the
impedance mismatch in an incremental fashion.

One derivative challenge to the first is the problem of simulating the
programmer's mistakes. Before a programmer can try to find mistakes, mistakes
have to be made. Technically the problem is the lack of a sufficiently large
available corpus of type errors for a meaningful experiment. Similar to
~\citet{lksfd-popl-2020}, we cover this gap by injecting bugs in typed programs
with mutation analysis~\cite{lipton1971fault, demillo1978hints,
jia2011analysis}. This results into a large number of buggy programs whose
mix-typed versions are suitable subjects for our experiment. However, mutation
analysis traditionally aims to inject bugs that are sophisticated enough to
challenge extensive test suites. As a consequence, the ill-typed mutants we care
about here are usually deemed \emph{incompetent} and mutation analysis
frameworks are fine-tuned to avoid them. To overcome this problem, we design a
set of mutators specific for type errors inspired by common errors when
programming with in Typed Racket.  These type errors cover a wide range of Typed
Racket's type system and moreover cause reliably runtime type errors in
mix-typed versions of mutants.

Another derivative challenge is the problem of mistakes in type specifications.
Gradual typing enables programmers to move an untyped code base into the typed
realm on an incremental basis. The retroactive addition of a specific type to a
component, even a library, may result from a misunderstanding of the code and
may thus be a mistake itself. Empirical evidence suggests that this scenario is
reasonably common~\cite{sta-nt-base-types, incorrect-ts, wmwz-ecoop-2017}, and
hence must be addressed by any evaluation method.  \ldots needs to also explore
the relation between blame and the location of bugs in case of wrong type
annotations. Similarly to the previous challenge, we deal with this one by
designing a set of mutators that ``break'' type annotations but leave programs
otherwise intact.

The second big challenge is the set of seemingly incomparable blame-assignment
strategies. This results in this paper focus on three points of the landscape:
the {\em natural\/} approach of Typed Racket~\cite{tf-dls-2006, tf-popl-2008,
tfffgksst-snapl-2017, tf-icfp-2010}, the {\em transient\/} one of Reticulated
Python~\cite{vsc-dls-2019,vss-popl-2017,vksb-dls-2014}, and the {\em optional\/}
typing approach of industrial systems. Each of these points comes with a
distinct blaming strategy and together they represent all proposed strategies so
far. In addition, instead of experimenting with the three approaches in
isolation, we have implemented transient and optional gradual typing as variants
of Typed Racket to seize the opportunity to compare them.  However,
implementations on top of the same platform are not enough for an
apples-to-apples comparison. The three approaches differ not only on their
blaming strategies but also on the type checks they perform. To untangle the
two, we supplement our experiment with two extra modes. The \emph{vanilla} mode
ignores blame and instead uses information from dynamic type error messages to
select how to proceed with migration. Hence it separates the effect of blame on
the experiment from that of the type checks. The \emph{control} mode randomly
selects which part of the program to migrate and serves as the null hypothesis
control for the overall methodology.
